\documentclass[11pt, twoside, openright]{report} % 
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[a4paper,width=150mm,top=25mm,bottom=25mm,bindingoffset=12mm]{geometry}
%\usepackage{biblatex}
\usepackage{setspace}
\usepackage[english]{babel} 
\usepackage{stmaryrd}
\usepackage{xcolor}
\usepackage{xspace}
%\usepackage{natbib}
\usepackage{csquotes}
\usepackage[style=apa, backend=biber]{biblatex}
%\DeclareLanguageMapping{american}{american-apa}
%\usepackage{apacite}
\addbibresource{references.bib}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{breqn}
% packages for reading results
\usepackage{pgfplotstable}
\usepackage{csvsimple}
\usepackage{siunitx}
\usepackage{lscape}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{adjustbox}
% define col width
\newcolumntype{Y}{>{\hsize=4\hsize}X}
\newcolumntype{s}{>{\hsize=0.25\hsize}X}
\graphicspath{ {images/} }

\linespread{1.25}
\counterwithout{footnote}{chapter}

\definecolor{Red}{RGB}{255,0,0}
\definecolor{Green}{RGB}{10,200,100}
\definecolor{Blue}{RGB}{10,100,200}
\definecolor{Orange}{RGB}{255,153,0}
\definecolor{Purple}{RGB}{139,0,139}

\newcommand{\denote}[1]{\mbox{ $[\![ #1 ]\!]$}}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand{\red}[1]{\textcolor{Red}{#1}}  
\newcommand{\eb}[1]{\textcolor{Blue}{[mht: #1]}}  
\newcommand{\mf}[1]{\textcolor{Orange}{[rl: #1]}}  
\newcommand{\pt}[1]{\textcolor{Purple}{[pt: #1]}} 

\DeclareMathOperator*{\E}{\mathbb{E}}
% define functions for reading results from csv
\newcommand{\datafoldername}{R4Tex}

% the following code defines the convenience functions
% as described in the main text below

% rlgetvalue returns whatever is the in cell of the CSV file
% be it string or number; it does not format anything
\newcommand{\rlgetvalue}[4]{\csvreader[filter strcmp={\mykey}{#3},
	late after line = {{,}\ }, late after last line = {{}}]
	{\datafoldername/#1}{#2=\mykey,#4=\myvalue}{\myvalue}}

% rlgetvariable is a shortcut for a specific CSV file (myvars.csv) in which
% individual variables that do not belong to a larger chunk can be stored
\newcommand{\rlgetvariable}[2]{\csvreader[]{\datafoldername/#1}{#2=\myvar}{\myvar}\xspace}

% rlnum format a decimal number
\newcommand{\rlnum}[2]{\num[output-decimal-marker={.},
	exponent-product = \cdot,
	round-mode=places,
	round-precision=#2,
	group-digits=false]{#1}}

\newcommand{\rlnumsci}[2]{\num[output-decimal-marker={.},
	scientific-notation = true,
	exponent-product = \cdot,
	round-mode=places,
	round-precision=#2,
	group-digits=false]{#1}}

\newcommand{\rlgetnum}[5]{\csvreader[filter strcmp={\mykey}{#3},
	late after line = {{,}\ }, late after last line = {{}}]
	{\datafoldername/#1}{#2=\mykey,#4=\myvalue}{\rlnum{\myvalue}{#5}}}

\newcommand{\rlgetnumsci}[5]{\csvreader[filter strcmp={\mykey}{#3},
	late after line = {{,}\ }, late after last line = {{}}]
	{\datafoldername/#1}{#2=\mykey,#4=\myvalue}{\rlnumsci{\myvalue}{#5}}}

% MH's command
\newcommand{\brmresults}[2]{\(\beta = \rlgetnum{#1}{Rowname}{#2}{Estimate}{3}\) (\rlgetnum{#1}{Rowname}{#2}{l.95..CI}{3}, \rlgetnum{#1}{Rowname}{#2}{u.95..CI}{3})}
%\brmresults{expt1_brm.csv}{condition}

\begin{document}
\input{title.tex}

\chapter*{Abstract}
Teaching artificial agents to communicate about the world in natural language in a way intuitive to humans is a difficult task. Rapid progress in deep learning has achieved impressive results in approximating this task by training deep neural agents on large annotated datasets to produce image captions. Yet only few attempts have been made to learn such communication from active interaction, akin to how humans learn language \parencite[e.~g.,][]{clark1996using}.
This thesis aims to take a step towards this task by training agents to produce referential messages in English based on interactive communication. More specifically, the agents learn to play a \emph{reference game} on real-world images within a multi-agent communication setting. Building on the work by \cite{lazaridou2020multi}, the thesis focuses on investigating the reasons for the so-called \emph{language drift} phenomenon---the deterioration of the agents' linguistic capabilities as they learn the task. Different possible causes are investigated experimentally, next to assessing several language drift metrics for capturing the phenomenon. More precisely, \emph{structural, semantic} and \emph{functional} language drifts are evaluated.
To this end, main experiments are conducted on the MS COCO dataset containing natural images \parencite{chen2015microsoft}; for comparison, additional experiments are conducted on the synthetic 3Dshapes dataset \parencite{burgess20183d}.
The experiments investigate whether the guaranteed presence of exhaustively descriptive captions in the dataset, the difficulty of the context, the co-adaptation potential of the agents and the strength of surface structure constraints on the agents' messages affect language drift. The results indicate that the difficulty of the visual context in which the game is learned, the co-adaptation potential of the agents as well as their architecture and the length of the annotations in the dataset have the largest impact on language drift. These results show that artificial agents can successfully learn to play a reference game in a realistic setting and do so in a context-dependent way, while calling for closer attention towards configurational aspects of multi-agent communication experiments.  


\chapter*{Acknowledgements}

I would like to thank Elia Bruni and Michael Franke for the opportunity to do this project and their guidance throughout the thesis. Their invaluable support and understanding allowed me to learn a great deal. I would also like to thank Fausto Carcassi and Berit Reise for their thoughtful feedback on the thesis.
Finally, I would like to thank my family, my partner Marcel Zaschka and my friends Berit Reise, Sarah J\"ahnichen, Lukas Schießer, Hedye Tayebi and Leonie Brodbeck for their moral and emotional support---I would not have been able to accomplish this work without them.

\tableofcontents
%\listoffigures
%\listoftables

\chapter{Introduction}
\label{chapter01}
\input{chapter01}

 \chapter{Technical Background}
 \label{chapter02}
 \input{chapter02}
 
\chapter{Related Work}
\label{chapter03}
\input{chapter03}

\chapter{Investigating Language Drift}
\label{chapter04}
\input{chapter04}

\chapter{Experiments}
\label{chapter05}
\input{chapter05}

\chapter{Discussion}
\label{chapter06}
\input{chapter06}

\chapter*{Declaration of Authorship}
I hereby certify that the work presented here is, to the best of my knowledge and belief, original and the result of my own investigations, except as acknowledged, and has not been submitted, either in part or whole, for a degree at this or any other university.

\vspace{2cm}
Signature:~\makebox[3in]{\hrulefill}

\vspace{1cm}
City, date:~\makebox[3in]{\hrulefill} 

\appendix
\chapter{Appendix}	
\label{appendix}
\input{appendix}

\printbibliography
%\bibliography{references}
\end{document}
